### 2017-05-06 2017mergingmerges/section/02-Processes.tex
Add a new section Breaking it Down to give a high level overview of how the system works

% TODO: Blend this in.
Note that we could construct the fused result machine in several ways. One option is to perform the case branch first and then pull from @sIn2@, another is to pull from @sIn2@ first and then perform the branch. By construction, the predicate used in the branch refers only to variables local to the @group@ machine, and the pull instruction from @B1@ stores its result in a variable local to the @merge@ machine. As the set of variables does not overlap, either ordering is correct. For this example we choose to perform the branch first, though will discuss the ramifications of this choice further in \S\ref{s:FusionOrder}.

@@

Overall, our fusion algorithm has taken two separate processes that we once imagined to be running concurrently, and has produced a single sequential result process that implements both. We have chosen a \emph{specific sequential order} in which to interleave the instructions that implement both source processes. As with Flow Fusion \cite{lippmeier2013data} we have performed the job of a concurrent scheduler at compile time. However, in contrast to Flow Fusion and similar systems, we do not need to organize statements into a fixed \emph{loop anatomy}, we simply merge them as they are. This allows us to implement a wider range of processes, including ones with nested loops that work on segmented streams, which we discuss further in \S\ref{s:FutureWork}.

To complete the implementation of our example from \S\ref{s:Introduction} we would now proceed to fuse the process from the final line (also a @group@) into this new result process. The order in which pairs of processes are fused together does matter, as does the order in which instructions are interleaved --- we discuss both points further in \S\ref{s:Evaluation}.

Finally, although the result process has a single shared heap, the bindings are guaranteed not to interfere. When we instantiated combinators to create the original source processes we introduced fresh names at that point. The stream buffer variables we additionally introduced for coordination were freshly created during fusion.

>>>>

@@

To complete the implementation of our example from \S\ref{s:Introduction} we would now proceed to fuse the process from the final line (also a @group@) into this new result process. The order in which pairs of processes are fused together does matter, as does the order in which instructions are interleaved --- we discuss both points further in \S\ref{s:Evaluation}. Although the result process has a single shared heap, the heap bindings from each fused process are guaranteed not to interfere. When we instantiated combinators to create the original source processes we introduced fresh names at that point.

% The stream buffer variables we additionally introduced for coordination were freshly created during fusion.

\subsection{Breaking It Down}
Overall, we started with a pure functional program as in \S\ref{s:Introduction}, reimagined it as a dataflow graph, then interleaved imperative code that implemented two of the operators in that dataflow graph. Importantly, we needed to ``break down'' the definition of each of the operators into imperative statements so that we could interleave their execution appropriately. We do this because the standard, single-threaded evaluation semantics of functional programs does not allow us to evaluate stream programs that contain both splits and joins in a space efficient way. Returning to the definition of @uniquesUnion@ from \S\ref{s:Introduction}, we cannot simply execute the @group@ operator on its entire input @sIn1@ before executing the @merge@ operator, as that would require us to buffer all data read from @sIn1@. Instead, in the fused result program we have effectively performed the job of a concurrent scheduler at compile time. The flow of control alternates between the implementation of each of the operators, but as the instructions are interleaved directly there is no overhead due to context switching.

The general approach of converting a pure functional program to a dataflow graph, then interleaving imperative code that implements the operators was also used in prior work on Flow Fusion \cite{lippmeier2013data}. However, in contrast to Flow Fusion and similar systems, with Machine Fusion we do not need to organize statements into a fixed \emph{loop anatomy} --- we simply merge them as they are. This allows us to implement a wider range of processes, including ones with nested loops that work on segmented streams, which we discuss in \S\ref{s:FutureWork}.

Note that trying to use lazy evaluation for an example such as @uniquesUnion@ does not solve the buffering problem. Suppose we converted each of the streams to lazy lists, and used definitions of @group@ and @merge@ that worked over these lists. As @uniquesUnion@ returns a pair of results, there is nothing preventing a consumer from demanding the first list (@sUnique@) in its entirity before demanding any of the elements from the second list (@sUnion@). If this were to happen then the runtime implementation would need to retain all elements of @sIn1@ before remanding any of @sIn2@, causing a space leak. Lazy evaluation is \emph{pull only} meaning that evaluation is driven by the consumer. The space efficiency of our fused program relies critically on the fact that operators can also \emph{push} their result values directly to their consumers, and that the consumers are not at liberty to defer handling of these values.

% TODO: Blend this in.
% Note that we could construct the fused result machine in several ways. One option is to perform the case branch first and then pull from @sIn2@, another is to pull from @sIn2@ first and then perform the branch. By construction, the predicate used in the branch refers only to variables local to the @group@ machine, and the pull instruction from @B1@ stores its result in a variable local to the @merge@ machine. As the set of variables does not overlap, either ordering is correct. For this example we choose to perform the branch first, though will discuss the ramifications of this choice further in \S\ref{s:FusionOrder}.
